---
title: "Transformation Elasticity for regional version of SIMPLE or GTAP or other Equilibrium Models derived by mrwater"
author: "Felicitas Beier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Setup, include = FALSE}
# R markdown set-up
knitr::opts_chunk$set(echo = TRUE)

# Data frame operations from tidyverse
library(tibble)

# mr universe functions
library(madrat)
library(magclass)
library(mrwater)

# For reading shape file 
library(sf)

```

```{r Settings0, include = FALSE}
version  <- "SIMPLEopt_2023_09_04"
year     <- "y2010"
scenario <- "ssp2"
efp      <- "off"

# Path
filepath       <- paste0(getwd(), "/Data/")
inputdatapath  <- paste0(getwd(), "/Data/", version, "/")
outputdatapath <- paste0(getwd(), "/Outputs/", version, "/")

```

## Introduction

Here, we use data from mrwater to derive the transformation elasticity between rainfed and irrigated land at an aggregated scale. 

The goal is to estimate a function for irrigated area depending on the yield gain through irrigation.
We know the technical maximum irrigation potential per region (maxIrrigArea), i.e. the potentially irrigated area (PIA) at a yield gain of zero. 
We want to fit a regression line to determine a continuous functional form that describes the relationship between PIA at different yield gains. 

The advantage of using mrwater is that it models PIA taking water availability and the upstream-downstream relationship of water flows as well as different water uses into account. Furthermore, it uses the information of gridded irrigated and rainfed yields as well as country-level crop prices to determine the yield gain through irrigation. This drives the allocation of areas that can potentially be irrigated. 

We derive for every grid cell several PIA at different yield gain thresholds. These can be summarized at the regional/country-/basin-level resolution in "PIA curves" (supply curves). These PIA curves that depict the maximum area that can potentially be irrigated at different yield gains. 
These is the modeled data that will be used to determine the relationship. 


```{r ReadData, include = FALSE, echo = FALSE}
# River structure for mapping to river basins
rs <- readRDS(system.file("extdata/riverstructure_stn_coord.rds",
                          package = "mrwater"))

# Economic irrigation area potential 
IrrigAreaEconCUR <- collapseNames(read.magpie(paste0(inputdatapath, "IrrigAreaEconCUR.mz"))[, , efp])

allCountries <- getItems(IrrigAreaEconCUR, dim = "iso")
allBasins    <- unique(rs$endcell)

# Yield gain range chosen for the analysis
gtrange <- getItems(IrrigAreaEconCUR, dim = 3)
```

```{r Shapefile, include = FALSE, echo = FALSE}
# Shape file provided:
nameOfShape <- "20230301_gtapv11"

map <- toolGetMappingCoord2Country(pretty = TRUE)
 
# Translate GTAP shape file to mapping to 67420 grid cells (0.5 degree)
unzip(paste0(nameOfShape, ".zip"), exdir = ".")

shpfile <- st_read(paste0(getwd(), "\\", nameOfShape, ".shp"))
head(shpfile, n = 4)
# # Plot it without converting geometry to x, y columns
# plot(st_geometry(shpfile))
# #plot(st_geometry(st_centroid(shpfile)), add = TRUE)
# 
# 
# a <- terra::vect(shpfile)
# b <- aggregate(a, 4, FUN = )
# 
# 
# terra::rasterize(x = sphfile, y = as.raster(IrrigAreaEconCUR))



```


```{r CreateDataFrames, include = FALSE, echo = FALSE}

# Transform raw data into data frame
baseData <- as.data.frame(as.array(IrrigAreaEconCUR))
baseData <- rownames_to_column(baseData, var = "Cell")
colnames(baseData) <- c("Cell", gtrange)
baseData <- reshape(baseData, direction = "long",
                    varying = list(names(baseData)[2:length(names(baseData))]),
                    idvar = "Cell",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")

# Create country data frame
ctryData <- as.data.frame(as.array(dimSums(IrrigAreaEconCUR, dim = c("x", "y"))))
ctryData <- rownames_to_column(ctryData, var = "iso")
colnames(ctryData) <- c("iso", gtrange)
ctryData <- reshape(ctryData, direction = "long",
                    varying = list(names(ctryData)[2:length(names(ctryData))]),
                    idvar = "iso",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")
ctryData$YieldGain <- as.integer(ctryData$YieldGain)


# Create basin data frame
tmp <- IrrigAreaEconCUR
getItems(tmp, dim = 1, raw = TRUE) <- paste(getItems(tmp, dim = 1), rs$endcell, sep = ".")
bsnData <- as.data.frame(as.array(dimSums(tmp, dim = c(1.1, 1.2, 1.3))))
bsnData <- rownames_to_column(bsnData, var = "basin")
colnames(bsnData) <- c("basin", gtrange)
bsnData <- reshape(bsnData, direction = "long",
                    varying = list(names(bsnData)[2:length(names(bsnData))]),
                    idvar = "basin",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")
bsnData$YieldGain <- as.integer(bsnData$YieldGain)


# Identify basins
basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))
# 3462 some river bains in Nevada; Great Basin?
# 2538 Columbia river basin? or Pacific Northwest Basin?

# tmp <- IrrigAreaEconCUR
# tmp[, , ] <- 0
# tmp[rs$coordinates[rs$endcell==3462], , ] <- 1

# Note: I could not find the Upper Snake River (have a look together with Iman)
  
```


The following graphs show the results for the Colorado river basin

```{r BasinExample, include = TRUE, echo = FALSE}

basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))

basins <- "4089"

for(b in basins) {
  
  data <- bsnData[bsnData$basin == b, ]
  plot(data$PIA, data$YieldGain,
       xlab = "Irrigated Area", ylab = "Yield Gain",
       xlim = c(min(data$PIA), max(data$PIA)),
       main = basinMap$Basin.Name[basinMap$Endcell == b])
}

```

and aggregate to the country-level for the United States

```{r CountryExample, include = TRUE, echo = FALSE}

countries <- "USA"

for(c in countries) {
  
  # create data frame for the selected country
  data <- ctryData[ctryData$iso == c,]
  
  # plot modeled data for the selected country
  plot(data$PIA, data$YieldGain,
       xlab = "Irrigated Area", ylab = "Yield Gain",
       xlim = c(0, max(data$PIA)),
       main = c)
}

```


# Basin Elasticity

I tried to fit the simulated data (little circles in the graph) to different functional forms using a linear regression model with the functional forms:

* y = a + b (1/x) (red line)
* y = a + b1 x + b2 x^2 + b3 x^3 (blue line)

The following graphs show the results for a selection of different river basins around the world. The show that the elasticities are heterogenous and different functional forms perform better in some river basins than in others.

```{r BasinHeterogeneity, include = TRUE, echo = FALSE}

basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))

basins <- c("4089",  # Colorado
            "24809", # Guadalquivir
            "29805", # Orange River
            "2538",  # Columbia River
            "9443",  # Mississippi
            "35804", # Nile
            "59064", # Chang Jian
            "50464", # Ganges
            "44955", # Indus
            "63145", # Murray 
            "58181", # Huang He
            "35257", # Danube
            "54662", # Mekong
            "18169", # Parana
            "43774", # Syr-Darya
            "43578") # Amu-Darya


for(b in basins) {
  
  data <- bsnData[bsnData$basin == b, ]
  
  # plot modeled data for the selected basin
  plot(data$PIA, data$YieldGain,
       xlab = "Irrigated Area", ylab = "Yield Gain",
       xlim = c(0, max(data$PIA)),
       main = c)
  
  # Fitting a 1/x function
  reg1 <- lm(YieldGain ~ I(1/PIA), data = data)
  curve(reg1$coefficients[1] + reg1$coefficients[2] * I(1/x),
        col = "red",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = 0, to = max(data$PIA), 
        add = TRUE)
  
  # Fitting a polynomial function
  reg2 <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
  curve(reg2$coefficients[1] + reg2$coefficients[2] * x +
        reg2$coefficients[3] * x^2 + reg2$coefficients[4] * x^3,
        col = "blue",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = 0, to = max(data$PIA),
        add = TRUE)
}

```


We find that the slope (elasticity) as well as the functional form vary by basin.

There are also completely inelastic basins (like Guadalquivir in Spain). For these, the function could not be fitted, for these the elasticity would be 0.


# Country elasticity

For country-level results, the functional form 1/x outperformed the polynomial.

```{r CountryHeterogeneity, include = TRUE, echo = FALSE}

countries <- c("DEU", "FRA", "ESP", "ITA",
               "USA", "MEX", "BRA",
               "IND", "CHN", "ZAF")

for(c in countries) {
  
  # create data frame for the selected country
  data <- ctryData[ctryData$iso == c,]
  
  # plot modeled data for the selected country
  plot(data$PIA, data$YieldGain,
       xlab = "Irrigated Area", ylab = "Yield Gain",
       xlim = c(0, max(data$PIA)),
       main = c)
  
  # Fitting a 1/x function
  reg1 <- lm(YieldGain ~ I(1/PIA), data = data)
  curve(reg1$coefficients[1] + reg1$coefficients[2] * I(1/x),
        col = "red",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = 0, to = max(data$PIA), 
        add = TRUE)
  
  # Fitting a polynomial function
  reg2 <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
  curve(reg2$coefficients[1] + reg2$coefficients[2] * x +
        reg2$coefficients[3] * x^2 + reg2$coefficients[4] * x^3,
        col = "blue",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = 0, to = max(data$PIA),
        add = TRUE)
}

```


# Determining the functional form 

The next step is to determine a functional form based on the data.

We would like to determine the transformation elasticity of rainfed to irrigated area based on the model data on PIAs at different yield gain thresholds. 

The goal is to estimate a function for irrigated area depending on the yield gain through irrigation.
We know the technical maximum irrigation potential per region (maxIrrigArea), i.e. the potentially irrigated area (PIA) at a yield gain of zero. 
We want to fit a regression line to determine a continuous functional form that describes the relationship between PIA at different yield gains. 

To do so, I would suggest to turn around the axes.
y = Potentially Irrigated Area (in Mha), and 
x = Irrigation yield gain (in USD/ha)

The technical potential (maxIrrigArea) is the intercept. 

The functional form would look like:

IrrigArea = MaxIrrigArea + f(YieldGain)

In the following section, I will try to find a suitable functional form and fit a regression using this form. 
I am considering to force the regression through the intersect. 

INSERT RESULTS HERE.

So far, I have tried around with functional forms like 1 / (value + x). 
I will try further functions here. 

```{r RegBasins, include = TRUE, echo = FALSE}
# basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))
# 
# basins <- c("4089",  # Colorado
#             "24809", # Guadalquivir
#             "29805", # Orange River
#             "2538",  # Columbia River
#             "9443",  # Mississippi
#             "35804", # Nile
#             "59064", # Chang Jian
#             "50464", # Ganges
#             "44955", # Indus
#             "63145", # Murray 
#             "58181", # Huang He
#             "35257", # Danube
#             "54662", # Mekong
#             "18169", # Parana
#             "43774", # Syr-Darya
#             "43578") # Amu-Darya
# 
# 
# for(b in basins) {
#   
#   data <- bsnData[bsnData$basin == b, ]
#   plot(data$YieldGain, data$PIA,
#        ylab = "Irrigated Area (in Mha)", xlab = "Yield Gain (in USD/ha)",
#        yaxs = "i", yaxs = "i",
#        ylim = c(min(data$PIA), max(data$PIA)),
#        main = basinMap$Basin.Name[basinMap$Endcell == b])
#   
#   # Technical maximum (PIWW at yield gain > 0)
#   techMax <- data$PIA[data$YieldGain == 0]
#   
#   reg <- lm(PIA ~ I(1 / ((1/techMax) + YieldGain)), data = data)
#   
#   curve(reg$coefficients[[1]] + reg$coefficients[[2]] * I(1 / ((1/techMax) + x)),
#         col = "pink",
#        ylab = "Irrigated Area (in Mha)", xlab = "Yield Gain (in USD/ha)",
#         from = min(data$YieldGain), to = max(data$YieldGain), 
#         add = TRUE)
#   
#   # Center data such that regression goes through y-axis intercept
#   data$PIA <- data$PIA - techMax
#   
#   reg <- lm(PIA ~ 0 + I(1 / ((1/techMax) + YieldGain)), data = data)
# 
#   # Transform data back
#   data$PIA <- data$PIA - techMax
#   
#   curve(techMax + reg$coefficients[[1]] * I(1 / ((1/techMax) + x)),
#         col = "red",
#        ylab = "Irrigated Area (in Mha)", xlab = "Yield Gain (in USD/ha)",
#         from = min(data$YieldGain), to = max(data$YieldGain), 
#         add = TRUE)
#   
#   data$YieldGain <- data$YieldGain + 0.00001
#   reg <- lm(PIA ~ I(1 / YieldGain), data = data)
#   
#     curve(reg$coefficients[[1]] + reg$coefficients[[2]] * I(1 / x),
#         col = "blue",
#        ylab = "Irrigated Area (in Mha)", xlab = "Yield Gain (in USD/ha)",
#         from = min(data$YieldGain), to = max(data$YieldGain), 
#         add = TRUE)
# 
#   
# }

```


# Create generic data for aggregated Equilibrium Models

For every country and every basin, we fit both functional forms and provide the parameters and model fit in a data table. 
These results can be used in equilibrium models that require information on the transformation elasticity of rainfed to irrigated land.

We report the country-level results in `countryElasticities.csv` and basin-level results in `basinElasticities.csv`.


```{r CountryResults, include = FALSE, echo = FALSE}

# countries <- allCountries
# 
# fct1 <- fct2 <- NULL
# for(c in countries) {
#   
#   # create data frame for the selected country
#   data <- ctryData[ctryData$iso == c,]
#   
#   # cut data when they stay the same for 3 consecutive years?
#   
#   
#   # Fitting a 1/x function
#   reg1 <- lm(YieldGain ~ I(1/PIA), data = data)
#   tmp1 <- data.frame(iso = as.character(c),
#                      adjR2_fct1 = summary(reg1)$adj.r.squared,
#                      a0_fct1 = reg1$coefficients[1],
#                      b1_fct1 = reg1$coefficients[2])
#   fct1 <- rbind(fct1, tmp1)
#   
#   # Fitting a polynomial function
#   reg2 <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
#   tmp2 <- data.frame(iso = as.character(c),
#                      adjR2_fct2 = summary(reg2)$adj.r.squared,
#                      a0_fct2 = reg2$coefficients[1],
#                      b1_fct2 = reg2$coefficients[2],
#                      b2_fct2 = reg2$coefficients[3],
#                      b3_fct2 = reg2$coefficients[4])
#   fct2 <- rbind(fct2, tmp2)
#   
#   # Combine to one data frame
#   countryOut <- merge(fct1, fct2)
#   # Save result in csv
#   write.table(countryOut, file = "countryElasticities.csv", sep = ",", dec = ".")
# }

```



```{r BasinResults, include = FALSE, echo = FALSE}

# basins <- allBasins

# Need to solve issue that it is not specified for all basins.

# fct1 <- fct2 <- NULL
# for(b in basins) {
#   
#   # create data frame for the selected basin
#   data <- bsnData[bsnData$basin == b,]
#   
#   # Fitting a 1/x function
#   if (!all(data$PIA == 0)) {
#      reg1 <- lm(YieldGain ~ I(1/PIA), data = data)
#      tmp1 <- data.frame(basin = as.character(b),
#                          adjR2_fct1 = summary(reg1)$adj.r.squared,
#                          a0_fct1 = reg1$coefficients[1],
#                          b1_fct1 = reg1$coefficients[2]) 
#   } else {
#      tmp1 <- data.frame(basin = as.character(b),
#                          adjR2_fct1 = 0,
#                          a0_fct1 = 0,
#                          b1_fct1 = 0) 
#   }
#   fct1 <- rbind(fct1, tmp1)
#   
#   # Fitting a polynomial function
#   if (!all(data$PIA == 0)) {
#     reg2 <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
#     tmp2 <- data.frame(basin = as.character(b),
#                        adjR2_fct2 = summary(reg2)$adj.r.squared,
#                        a0_fct2 = reg2$coefficients[1],
#                        b1_fct2 = reg2$coefficients[2],
#                        b2_fct2 = reg2$coefficients[3],
#                        b3_fct2 = reg2$coefficients[4])
#   } else {
#     tmp2 <- data.frame(basin = as.character(b),
#                        adjR2_fct2 = 0,
#                        a0_fct2 = 0,
#                        b1_fct2 = 0,
#                        b2_fct2 = 0,
#                        b3_fct2 = 0)
#   }
#   fct2 <- rbind(fct2, tmp2)
#   
#   # Combine to one data frame
#   basinOut <- merge(fct1, fct2)
#   # Save result in csv
#   write.table(basinOut, file = "basinElasticities.csv", sep = ",", dec = ".")
# }

```


# Conclusion 

Main message: don't use a global uniform function and value for the transformation elasticity; better to use modeled data to derive a country- or basin-level relationship and inform the model


Remaining questions:

* How to choose a good data range for the fit generically?
* Does the fit y = yield gain, x = irrigated area make sense or should it be the other way around? 
* Is there a way to force the function to meet the correct max value (intercept with x with the last point rather than using the intercept parameter estimated?)
