---
title: "Transformation Elasticity for regional version of SIMPLE derived by mrwater"
author: "Felicitas Beier"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r Setup, include = FALSE}
# R markdown set-up
knitr::opts_chunk$set(echo = TRUE)

# Data frame operations from tidyverse
library(tibble)

# mr universe functions
library(madrat)
library(magclass)
library(mrwater)
```

```{r Settings0, include = FALSE}
version  <- "SIMPLEopt_2023_09_04"
year     <- "y2010"
scenario <- "ssp2"
efp      <- "off"

# Path
filepath       <- paste0(getwd(), "/Data/")
inputdatapath  <- paste0(getwd(), "/Data/", version, "/")
outputdatapath <- paste0(getwd(), "/Outputs/", version, "/")

```

## Introduction
Here, we assess use data from mrwater to derive the transformation elasticity between rainfed and irrigated land at an aggregated scale. 

```{r ReadData, include = FALSE, echo = FALSE}
# River structure
rs <- readRDS(system.file("extdata/riverstructure_stn_coord.rds",
                          package = "mrwater"))

# Economic irrigation area potential 
IrrigAreaEconCUR <- collapseNames(read.magpie(paste0(inputdatapath, "IrrigAreaEconCUR.mz"))[, , efp])

# Yield gain range chosen for the analysis
gtrange <- getItems(IrrigAreaEconCUR, dim = 3)
```


```{r CreateDataFrames, include = FALSE, echo = FALSE}

# Transform raw data into data frame
baseData <- as.data.frame(as.array(IrrigAreaEconCUR))
baseData <- rownames_to_column(baseData, var = "Cell")
colnames(baseData) <- c("Cell", gtrange)
baseData <- reshape(baseData, direction = "long",
                    varying = list(names(baseData)[2:length(names(baseData))]),
                    idvar = "Cell",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")

# Create country data frame
ctryData <- as.data.frame(as.array(dimSums(IrrigAreaEconCUR, dim = c("x", "y"))))
ctryData <- rownames_to_column(ctryData, var = "iso")
colnames(ctryData) <- c("iso", gtrange)
ctryData <- reshape(ctryData, direction = "long",
                    varying = list(names(ctryData)[2:length(names(ctryData))]),
                    idvar = "iso",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")
ctryData$YieldGain <- as.integer(ctryData$YieldGain)


# Create basin data frame
tmp <- IrrigAreaEconCUR
getItems(tmp, dim = 1, raw = TRUE) <- paste(getItems(tmp, dim = 1), rs$endcell, sep = ".")
bsnData <- as.data.frame(as.array(dimSums(tmp, dim = c(1.1, 1.2, 1.3))))
bsnData <- rownames_to_column(bsnData, var = "basin")
colnames(bsnData) <- c("basin", gtrange)
bsnData <- reshape(bsnData, direction = "long",
                    varying = list(names(bsnData)[2:length(names(bsnData))]),
                    idvar = "basin",
                    timevar = "YieldGain", times = gtrange,
                    v.names = "PIA")
bsnData$YieldGain <- as.integer(bsnData$YieldGain)


# Identify basins
basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))
# 3462 some river bains in Nevada; Great Basin?
# 2538 Columbia river basin? or Pacific Northwest Basin?

# tmp <- IrrigAreaEconCUR
# tmp[, , ] <- 0
# tmp[rs$coordinates[rs$endcell==3462], , ] <- 1

# Note: I could not find the Upper Snake River (have a look together with Iman)
  
```

# Basin Elasticity

I tried to fit the simulated data (little circles in the graph) to some functional form using a regression. 
I tried 1/x (red) and a polynomial function (blue). 

```{r BasinHeterogeneity, include = TRUE, echo = FALSE}

basinMap <- read.csv(system.file("extdata/RiverBasinMapping.csv", package = "mrwater"))

basins <- c("4089",  # Colorado
  "24809", # Guadalquivir
  "29805", # Orange River
  "2538",  # Columbia River
  "9443",  # Mississippi
  "35804", # Nile
  "59064", # Chang Jian
  "50464", # Ganges
  "44955", # Indus
  "63145", # Murray 
  "58181", # Huang He
  "35257", # Danube
  "54662", # Mekong
  "18169", # Parana
  "43774", # Syr-Darya
  "43578") # Amu-Darya


for(b in basins) {
  data <- bsnData[bsnData$basin == b,]
  plot(data$PIA, data$YieldGain,
     xlab = "Irrigated Area", ylab = "Yield Gain",
     xlim = c(min(data$PIA), max(data$PIA)),
     main = basinMap$Basin.Name[basinMap$Endcell==b]
     )
  
  # Fitting a 1/x function
  reg <- lm(YieldGain ~ I(1/PIA), data = data)
  curve(reg$coefficients[1] + reg$coefficients[2] * I(1/x),
        col = "red",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = min(data$PIA), to = max(data$PIA), 
        add = TRUE)
  
  # Fitting a polynomial function
  reg <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
  curve(reg$coefficients[1] + reg$coefficients[2] * x + 
        reg$coefficients[3] * x^2 + reg$coefficients[4] * x^3,
      col = "blue",
      xlab = "Irrigated Area", ylab = "Yield Gain",
      from = min(data$PIA), to = max(data$PIA), 
      add = TRUE)
}

```

Note: I also tried log(x), sqrt(x), x + x^2. They all look worse.

The functional form seems to vary by basin, wich I believe is a bit problematic. 
There are also completely inelastic basins (like Guadalquivir in Spain). For these, the function could not be fitted, but I think these are special cases for which we just assume completely inelastic demand (elasticity is 0).

Questions:
How to choose a good data range? (probably country-specific)
How to choose a good functional form? (seems to be different by country... we need a flexible functional form)

# Country elasticity 

I repeated this exercise for some countries. In that case the functional form 1/x seemed to fit better.
I still wonder whether this is good enough (see IND, CHN, USA). 

```{r CountryHeterogeneity, include = TRUE, echo = FALSE}

countries <- c("DEU", "FRA", "ESP", "ITA",
               "USA", "MEX", "BRA",
               "IND", "CHN", "ZAF")

for(c in countries) {
  data <- ctryData[ctryData$iso == c,]
  plot(data$PIA, data$YieldGain,
     xlab = "Irrigated Area", ylab = "Yield Gain",
     xlim = c(0, max(data$PIA)),
     main = c)
  
  # Fitting a 1/x function
  reg <- lm(YieldGain ~ I(1/PIA), data = data)
  curve(reg$coefficients[1] + reg$coefficients[2] * I(1/x),
        col = "red",
        xlab = "Irrigated Area", ylab = "Yield Gain",
        from = 0, to = max(data$PIA), 
        add = TRUE)
  
  # Fitting a polynomial function
  # reg <- lm(YieldGain ~ PIA + I(PIA^2) + I(PIA^3), data = data)
  # curve(reg$coefficients[1] + reg$coefficients[2] * x + 
  #       reg$coefficients[3] * x^2 + reg$coefficients[4] * x^3,
  #     col = "blue",
  #     xlab = "Irrigated Area", ylab = "Yield Gain",
  #     from = 0, to = max(data$PIA), 
  #     add = TRUE)
}


```

Do you have better ideas how to approach this exercise? 
Does the fit y = yield gain, x = irrigated area make sense or should it be the other way around? 
